Estimating duration using survival analysis

### 2024-03-22 cis-perfect-testing.tex
finish 5.3 rewrite

The CIS sampling is less frequent, but with has a large sample size (\numprint{4800} detected infection episodes) and a long follow-up.
Therefore, it is suitable for estimating the tail of the duration distribution.
In this chapter, I develop a method to estimate the duration distribution using CIS data, assuming no false negatives or positives.

In \cref{perf-test:sec:problem}, I describe the challenges analysing the data presents.
This motivates using a survival analysis framework which I introduce in \cref{perf-test:sec:surv-analysis}, before deriving the statistical model for the CIS setting in \cref{perf-test:sec:model}.
Next, in \cref{perf-test:sec:parameters-priors}, I consider how to parameterise the survival function and choose priors; these include vague and smoothing priors as well as including information from \cref{E-ATACCC}.
I evaluate the model and priors using a simulation study (\cref{perf-test:sec:simulation-study}), showing that they perform well.
Finally, I discuss the results in \cref{perf-test:sec:discussion}.

@@

Infection episodes in an individual's one week testing period are over-represented in the detected infections.
This is because more frequent testing will detect a higher proportion of infections.
Therefore, for the average detected episode, $p(\text{episode $j$ detected} \mid b_j \in [l_j^{(b)}, r_j^{(b)}]) > p(\text{episode $j$ detected})$.
Therefore, conditioning on $b_j \in [l_j^{(b)}, r_j^{(b)}]$ is inappropriate.

@@

In this section, I develop a statistical model accounting for double interval censoring and undetected episodes.
This approach extends the framework of \textcite{heiseyModelling} to include all individuals in the CIS, including those without detected episodes.

The number of missed episodes is now a latent quantity that needs to be considered.
Marginalising over the number of missed episodes is possible analytically.
This makes inference tractable using the large sample size and the computing resources available within the SRS (see \cref{E-biology-data:sec:SRS}).
In particular, the posterior can be expressed within modern, high-performance Bayesian software such as Stan.

The model will be used to analyse the 4,800 episodes first detected between 10th Oct 2020 and 6th Dec 2020 inclusive, with negatives bounding the start and end time of the episode.
This period is denoted as $t = 1, \dots, T$.
A negative test before the start of the episode provides a lower bound on the episode's start time; otherwise, there is little information on its length.
Similar information comes from the negative test after the end of the episode.
This is typically observed, unless the individual is lost to follow-up before their episode is ended, which happens rarely.

@@

Let the total number of infection episodes in the cohort be $\ntot = \ndet + \nnodet$ where $\ndet$ is the number of detected infections and $\nnodet$ is the number of undetected infections.
Let $\Ncis$ be the number of individuals in the cohort.
%Without loss of generality, let the individuals with a detected episode be indexed by $i = 1, \dots, \ndet$ and the individuals without a detected episode be indexed by $i = \ndet + 1, \dots, \Ncis$.
Let $i = 1, \dots, \Ncis$ index individuals and $j = 1, \dots \ntot$ index infection episodes.
The test schedule $\sched_i$ is the set of times individual $i$ is tested at, from their last test prior to time 1 (if it exists) up until time $T$.
This is all the times the infection episode could be detected at and be included, plus one prior test.
For example, for the individual in \cref{perf-test:sec:undetected}, $\sched_i = \{ 0, 7, 14, 21, 28, 56 \}$.
Each individual has exactly one test schedule, but a test schedule could have multiple individuals associated with it.

All the relevant information about the infection episode $j$ can be fully characterized by the (unobserved) triplet $(b_j, e_j, i(j))$ where $b_j$ and $e_j$ are the times the episode begin and end respectively and $i(j)$ denotes the individual the episode occurred in.
The episode's duration is $d_j = e_j - b_j + 1$ (see \cref{E-inc-prev:sec:definitions}).
%, the number of days the individual in which the episode occurred, was positive for 
% This triplet is assumed iid.

% The times $b_j$ and $e_j$ span possible infection and recovery times.
% $\sched_j$ is one of the observed test schedules, with any other test schedule assumed to occur with probability zero.

Over the short calendar period considered, each individual has zero or one detected episodes.
Let $i \in \set{D}$ indicate individual $i$ has a detected episode.
I define the precise conditions for this to occur in \cref{perf-test:sec:prob-undetected}.
% That is, $i \in \set{D}$ if and only if there exists a $j$ such that $b_{j}$ and $e_{j}$ such that $(b_{j}, e_{j}, \sched_i)$ is a detected episode.

To relate the possible value for $b_j$ and $e_j$, given a test schedule $\sched_{i(j)}$, to the data, I define three subsets partitioning the space of possible infection and recovery times (\ie: are strict subsets of $T \times T$).
Figure \ref{perf-test:fig:partitionSpace} shows this graphically.
Each episode, whether detected or not, must fall into one of these three classes with a probability that is a function of the parameters of the survival function, $\vec{\vec{\theta}}$, discussed in \cref{perf-test:sec:parameters-priors}.

\begin{itemize}
\item
  Admissible episodes, $\alpha_i$, which have an infection and recovery time consistent with the data observed.
  This space is empty for $i \notin \set{D}$; in this case individual $i$ had $n_{ia} = 0$ admissible episodes.
  Otherwise, if $i \in \set{D}$, then individual $i$ had $n_{ia} =1$ admissible episode.
  Conditional on the episode occurring in individual $i$, the episode is admissible with probability $p_{ia} = \prob((B_j, E_j) \in \alpha_i \mid \vec{\theta}, i(j) = i)$.
  \label{perf-test:def:admissible}
\item
  Undetected episodes, $\Omega_i$, which have an infection and recovery time such that they would not have tested positive.
  An unknown number, $n_{iu}$ of these occur in individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in individuals with test schedule $\sched_i$, the episode is undetected with probability $p_{iu} = \prob((B_j, E_j) \in \Omega_i \mid \vec{\theta}, i(j) = i)$.
\item
  Inadmissible episodes, $\beta_i$, which are not consistent with the data observed but would have been
  detected.
  These are all remaining episodes not in the previous sets, with $n_{i\bar{a}} = 0$ of these occurred in any individuals with test schedule $\sched_i$.
  Conditional on the episode occurring in these individuals, the episode inadmissible with probability $p_{i\bar{a}} = \prob((B_j, E_j) \in \beta_i \mid \vec{\theta}, i(j) = i)$.
\end{itemize}

\begin{figure}
\includegraphics[width=\textwidth]{cis-perfect-testing/regions_diag}
\caption[Admissible, inadmissible, and undetected infections]{The regions (as defined in the main text) for an individual with test schedule $\sched_i$ whose first test was at time 0 and negative, with subsequent
negative tests at times 7, 14, and 56, and positive tests at times 21
and 28. Inadmissible regions are unshaded. The black (impossible) region covers the cases in which recovery occurs prior to infection. \label{perf-test:fig:partitionSpace}}
\end{figure}

@@

Each episode occurs in one of the classes independently and with the same probability.
Therefore, conditional on $\ntot$, the number of episodes in each class is multinomial.
Specifically:
\begin{align}
  \begin{bmatrix}
    n_{1a} \\ n_{1u} \\ n_{1\bar{a}} \\ \vdots \\ n_{\Ncis,a} \\ n_{\Ncis,u} \\ n_{\Ncis,\bar{a}}
  \end{bmatrix}
  \mid \ntot
  =
  \MN\left(
    \ntot,
    \frac{1}{\Ncis}
    \begin{bmatrix}
      p_{1a} \\ p_{1u} \\ p_{1\bar{a}} \\ \vdots \\ p_{\Ncis,a} \\ p_{\Ncis,u} \\ p_{\Ncis,\bar{a}}
    \end{bmatrix}
  \right)
\end{align}

@@

\subsection{Posterior density}

Let $\na = (n_{1a}, \dots, n_{\Ncis a})^T$ be the vector of counts of episodes in the admissible classes.
This represents all our observations.
Therefore, the posterior density we are interested in is $p(\vec{\theta} \mid \na)$.
It can be written as follows.
\begin{align}
p(\vec{\theta} \mid \na)
&\propto p(\vec{\theta}) p(\na \mid \vec{\theta}) \\
&= p(\vec{\theta}) \sum_{\ntot=\ndet}^\infty p(\ntot \mid \vec{\theta}) p(\na \mid \ntot, \vec{\theta})
\label{perf-test:eq:posterior1}
\end{align}
where the last line augments the likelihood with the total number of episodes, $\ntot$.

From \cref{perf-test:sec:episode-types}, counts $\na$ and $\nnodet$ conditional on $\ntot$ are jointly distributed multinomially.
All $n_{i\bar{a}} = 0$, therefore these classes can be ignored; counts of zero do not contribute to the multinomial probability mass function (they multiply by 1).
The support of the distribution requires $\nnodet = \ntot - \sum_{i=1}^{\Ncis} n_{ia} = \ntot - \ndet$ and $\ntot \geq \ndet$.
Hence:
\begin{align}
p(\na \mid \ntot, \vec{\theta})
&= p(\nnodet  = \ntot - \na, \na\mid \ntot, \vec{\theta}) \\
&= \frac{\ntot!}{(\ntot - \ndet)! \prod_{i=1}^{\Ncis} n_{ia}!}  \pnodet^{\ntot - \ndet}\prod_{i=1}^{\Ncis} \left(\frac{1}{\Ncis}p_{ia}\right)^{n_{ia}} \\
&\propto \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \label{perf-test:eq:augmented-likelihood}
\end{align}
the final line follows because $n_{ia} = 1$ for $i \in \set{D}$ and $n_{ia} = 0$ otherwise.

Substituting \cref{perf-test:eq:augmented-likelihood} into \cref{perf-test:eq:posterior1}:
\begin{align}
&\propto p(\vec{\theta}) \sum_{\ntot=\ndet}^\infty \left( p(\ntot \mid \vec{\theta}) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \right) \\
&= p(\vec{\theta}) \left( \prod_{i \in \set{D}} p_{ia} \right) \left( \sum_{\ntot=\ndet}^\infty p(\ntot \mid \vec{\theta}) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \right) \label{perf-test:eq:posterior2}.
\end{align}
The rest of this section derives expressions for each $p_{ia}$, $p_{u}$ and the sum.

\subsection{Expression for the sum}

To start, I derive an analytical solution to $\sum_{\ntot=\ndet}^\infty p(\ntot \mid \vec{\theta}) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet}$ under the prior $\ntot \dist \NBc(\mu, r)$.
I assume that $\vec{\theta}$, which are only the parameters of the survival distribution, has a prior independent of $\ntot$, the number of infections that occurred in the cohort.
Therefore, $p(\ntot \mid \vec{\theta}) = p(\ntot)$.
Putting a negative binomial prior on $\ntot$ is equivalent to the following gamma-Poisson composite; its use simplifies the derivation.
\begin{align}
\ntot \mid \lambda &\dist \Poi(\lambda) \\
\lambda &\dist \GamDist(a, b)
\end{align}
where $b = r / \mu$ and $a = r$.
Hence:
\begin{align}
&\sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} p(\ntot) \\
&= \int \sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} p(\ntot \mid \lambda) p(\lambda) d\lambda &\text{$\lambda$ explicit}\\
&= \int \sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} \frac{\lambda^{\ntot} e^{-\lambda}}{\ntot!} p(\lambda) d\lambda &\ntot \dist \Poi\\
%&= \int \sum_{\ntot=\ndet}^\infty \frac{1}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} \lambda^{\ntot-\ndet} \lambda^{\ndet} e^{-\lambda} p(\lambda) d\lambda \\
&= \int \lambda^{\ndet} e^{-\lambda} p(\lambda) \sum_{\nnodet=0}^\infty \frac{(\pnodet \lambda)^{\nnodet}}{\nnodet!} d\lambda &\nnodet = \ntot-\ndet\\
&= \int \lambda^{\ndet} e^{-\lambda} p(\lambda) e^{\lambda \pnodet} d\lambda &\text{Maclaurin series of $e$} \\
&= \int \lambda^{\ndet} e^{-\lambda(1 - \pnodet)} \frac{b^a}{\Gamma(a)} \lambda^{a-1} e^{-b\lambda} \lambda d\lambda &\text{Gamma pdf}\\
&= \int \frac{b^a}{\Gamma(a)} \lambda^{a+\ndet-1} e^{-(b+1-\pnodet)\lambda} \lambda d\lambda \\
&= \frac{b^a}{\Gamma(a)} \frac{\Gamma(a+\ndet)}{(b+1-\pnodet)^{a+\ndet}} &\text{Gamma pdf}\\
&\propto (b+1-\pnodet)^{-(a+\ndet)} \\
&= (r/\mu + 1 - \pnodet)^{-(r+\ndet)} &\text{sub in $\mu$ and $r$}\\
&\propto(r + \mu (1- \pnodet))^{-(r+\ndet)}.
\end{align}

@@

An episode $j$ in individual $i(j)$ is detected if and only if all the following conditions are met.
\begin{enumerate}
    \item $\exists t \in \sched_{i(j)}$ such that $b_j \leq t \leq e_j$; this condition is equivalent to having at least one positive test for the episode.
    \item $\exists t \in \sched_{i(j)}$ such that $t < b_j$.
      For individuals enrolled during the period considered ($\min \sched_{i(j)} > 0$), this ensures that the beginning of the episode is lower bounded.
      For individuals enrolled prior to the period considered ($\min \sched_{i(j)} \leq 0$), this means that the episode was not detected prior to time 1.
    \item $\exists t \in \sched_{i(j)}$ such that $t > e_j$, upper bounding the end of the episode.
      For episodes detected in the period I consider, this happens a negligible number of times, only in individuals lost to follow-up; therefore, I neglect this possibility.
      For a new context, including recent infections, this condition could be relaxed by considering episodes that do not meet this criteria as right censored.
\end{enumerate}

% An episode is undetected if and only if no tests are performed during the episode or if there was no negative test prior to the episode.
% Equivalently, that the first test at or after $b$ is after $e$, or that there is no negative test prior to $b$.
To formalise this, first define $\tau_{\sched_i}(t)$ as the time until the next test at or after time $t$ in the schedule $\sched_i$:
\begin{align}
\tau_{\sched_i}(t) &= \min \{ t' \in \sched_i : t' \geq t \} - t.
\end{align}
The condition that $\exists t \in \sched_{i(j)}$ such that $b_j \leq t \leq e_j$ can now be expressed as $e_j \geq b_j + \tau_{\sched_{i(j)}}(b_j)$.
Equivalently, $d_j \geq \tau_{\sched_{i(j)}}(b_j) + 1$.

% Then $\Omega_i$ can be written as:
% \begin{align}
% \Omega_i = \{ (b, e) \ssep \tau_{\sched_i}(b) + b > e \vee b \leq \min(\sched_i) \}.
% \end{align}
Therefore, suppressing the conditioning on $\vec{\theta}$ and $i(j) = i$:
\begin{align}
1 - p_{iu}
&= 1 - \prob((B_{j}, E_{j}) \in \Omega_i) \\
&= \prob(D_j \geq \tau_{\sched_{i}}(B_j)+ 1, B_j > \min \sched_{j}) \\
&= \sum_{b = \min \sched_{i} + 1}^T \prob(D_j \geq \tau_{\sched_{i}}(b) + 1 \mid B_j = b) \prob(B_j = b)\\
&= \frac{1}{T} \sum_{b = \min \sched_{i} + 1}^T S_{\vec{\theta}}(\tau_{\sched_{i}}(b) + 1).
\intertext{Substituting into \cref{perf-test:eq:pu}:}
1 - p_u
& \propto \sum_{i=1}^{\Ncis} \sum_{b = \min \sched_{i} + 1}^T S_{\vec{\theta}}(\tau_{\sched_{i}}(b) + 1).
\end{align}

>>>>

The CIS sampling is less frequent, but with has a large sample size (\numprint{4800} detected infection episodes) and a long follow-up.
Therefore, it is suitable for estimating the tail of the duration distribution.
In this chapter, I develop a method to estimate the duration distribution using CIS data, assuming no false negatives or positives.
Formally, I aim to estimate the survival function $S$ defined in \cref{inc-prev:sec:definitions}.

In \cref{perf-test:sec:problem}, I describe the challenges analysing the data presents.
This motivates using a survival analysis framework which I introduce in \cref{perf-test:sec:surv-analysis}, before deriving the statistical model for the CIS setting in \cref{perf-test:sec:model}.
Next, in \cref{perf-test:sec:parameters-priors}, I consider how to parameterize the survival function and choose priors; these include vague and smoothing priors as well as including information from \cref{E-ATACCC}.
I evaluate the model and priors using a simulation study (\cref{perf-test:sec:simulation-study}), showing that they perform well.
Finally, I discuss the results in \cref{perf-test:sec:discussion}.

@@

Infection episodes in an individual's one-week testing period are over-represented in the detected infections.
This is because more frequent testing will detect a higher proportion of infections.
Therefore, for the average detected episode, $p(\text{episode $j$ detected} \mid b_j \in [l_j^{(b)}, r_j^{(b)}]) > p(\text{episode $j$ detected})$.
Therefore, conditioning on $b_j \in [l_j^{(b)}, r_j^{(b)}]$ is inappropriate.

@@

In this section, I develop a statistical model accounting for double interval censoring and undetected episodes.
In \cref{perf-test:sec:data-notation}, I introduce the data and notation, defining which quantities are observed and which are latent.
Then, in \cref{perf-test:sec:episode-types}, I apply the framework of \textcite{heiseyModelling} to classify possible episode beginning and end types into \emph{admissible} episodes, which match the doubly interval censored pattern of the observations; \emph{undetected} episodes, which may have occurred but would have been undetected; and \emph{inadmissible} episodes, which the data rules out having occurred.
The number of such episodes in each class is a random variable, with a multinomial distribution, conditional on the latent total number of infections (including undetected infections).

In \cref{perf-test:sec:posterior-density}, I outline the form of the posterior density.
The following subsections then derive its various components.
Unlike in \textcite{heiseyModelling}, I must consider the probability of undetected episodes for each individual seperately because they have differing visit times.
In addition, I include the information from the individuals without a detected episode.

Unlike prior work, I marginalise the likelihood over the number of undetected episodes.
This makes inference tractable using the large sample size and the computing resources available within the SRS (see \cref{E-biology-data:sec:SRS}).
In particular, the posterior can be expressed within modern, high-performance Bayesian software such as Stan.

\subsection{Data and notation} \label{perf-test:sec:data-notation}

I analyse the 4,800 episodes first detected between 10th Oct 2020 and 6th Dec 2020 inclusive, with negatives bounding the start and end time of the episode.
This period is denoted as $t = 1, \dots, T$.
A negative test before the start of the episode provides a lower bound on the episode's start time; otherwise, there is little information on its length.
Similar information comes from the negative test after the end of the episode.
A negligible number of episodes have a negative test after the end of the episode.
I refer to these episodes as \emph{detected}; in \cref{perf-test:sec:prob-undetected} I formalise the conditions for an episode to be detected.
Let $\ndet$ denote the total number of detected infections.

The cohort of individuals considered is all those in the CIS who could have had a detected test.
This requires that they had at least three tests before time $T$ (a negative prior to the episode, a positive, and a negative following the episode).
Index these individuals by $i = 1, \dots, \Ncis$ where $\Ncis$ is the total number of such individuals.

An individual $i$ has a \emph{test schedule} $\sched_i$.
The test schedule $\sched_i$ is the set of times individual $i$ is tested at, from their last test prior to time 1 (if it exists) up until time $T$.
This is all the times the infection episode could be detected at and be included, plus one prior test.
For example, for the individual in \cref{perf-test:sec:undetected}, $\sched_i = \{ 0, 7, 14, 21, 28, 56 \}$.
Each individual has exactly one test schedule, but it is possible that $\sched_i = \sched_{i'}$ for $i \neq i'$.
I assume that the test schedules are uninformative on all quantities of interest because they are part of the study design.
Therefore, I condition on them implicitly in all probabilities that follow.

For a detected episode $j$, five pieces of data are observed (see \cref{E-biology-data:sec:cis-episodes} for how these are constructed).
The individual in which the episode occurs, $i(j)$; the earliest and latest time the episode could have begun, $l_j^{(b)}$ and $r_j^{(b)}$ respectively; and the earliest and latest time the episode could have ended, $l_j^{(e)}$ and $r_j^{(e)}$ respectively.
Over the short calendar period considered, each individual has zero or one detected episodes.
Let $i \in \set{D}$ indicate individual $i$ has a detected episode.

In addition, the individuals in the cohort experienced an unknown number of undetected infections, $\nnodet$.
The total number of infections in the cohort is $\ntot = \ndet + \nnodet$ and I index them with $j = 1, \dots, \ntot$.

Any infection episode, whether detected or not, can be fully characterised by the triplet $(b_j, e_j, i(j))$ where $b_j$ and $e_j$ are the times the episode begin and end respectively, and $i(j)$ denotes the individual the episode occurred in.
The episode's duration is $d_j = e_j - b_j + 1$ (see \cref{E-inc-prev:sec:definitions}).
For all $j$, the exact values $b_j$ and $e_j$ are latent.
If $j$ is undetected then $i(j)$ is also latent and nothing is known about $b_j$ or $e_j$.
If $j$ is detected, then $i(j)$ is known; $b_j \in [l_j^{(b)}, r_j^{(b)}]$; and $e_j \in [l_j^{(e)}, r_j^{(e)}]$.
These latter two properties are what makes the data doubly interval censored.

@@

To relate the possible value for $b_j$ and $e_j$, given a test schedule $\sched_{i(j)}$, to the data, I define three subsets partitioning the space of possible infection and recovery times (\ie: are strict subsets of $T \times T$).
Figure \ref{perf-test:fig:partitionSpace} shows this graphically.
Each episode, whether detected or not, must fall into one of these three classes with a probability that is a function of the parameters of the survival function, $\vec{\vec{\theta}}$, specified in \cref{perf-test:sec:parameters-priors}.

\begin{itemize}
\item
  Admissible episodes, $\alpha_i$, which have an infection and recovery time consistent with the data observed.
  This space is empty for $i \notin \set{D}$; in this case individual $i$ had $n_{ia} = 0$ admissible episodes.
  Otherwise, if $i \in \set{D}$, then individual $i$ had $n_{ia} =1$ admissible episode.
  The admissible region is defined by the intersection of the intervals $[l_j^{(b)}, r_j^{(b)}]$ and $[l_j^{(e)}, r_j^{(e)}]$.
  This is the purple area in \cref{perf-test:fig:partitionSpace}; the boundaries in the $x$ direction of the area are $l_j^{(b)}$ and $r_j^{(b)}$ and in the $y$ direction are $l_j^{(e)}$ and $r_j^{(e)}$.
  The shape of this region accounts for the doubly interval censored nature of the data.
  Conditional on the episode occurring in individual $i$, the episode is admissible with probability $p_{ia} = \prob((B_j, E_j) \in \alpha_i \mid \vec{\theta}, i(j) = i)$.
  \label{perf-test:def:admissible}
\item
  Undetected episodes, $\Omega_i$, which have an infection and recovery time such that they would not have tested positive.
  An unknown number, $n_{iu}$ of these occur in individual $i$.
  The undetected region is the red region in \cref{perf-test:fig:partitionSpace}, which are when the individual is infected then recovers before their next test, plus when the individual is infected before their enrolment.
  The shape of this region accounts for the pattern of possible undetected infections.
  Conditional on the episode occurring in individual $i$, the episode is undetected with probability $p_{iu} = \prob((B_j, E_j) \in \Omega_i \mid \vec{\theta}, i(j) = i)$.
\item
  Inadmissible episodes, $\beta_i$, which would be detected (if they occur) but were not.
  Therefore, we know $n_{i\bar{a}} = 0$ of these occurred in individual $i$.
  These are all remaining episodes not in the previous sets.
  Conditional on the episode occurring in these individuals, the episode inadmissible with probability $p_{i\bar{a}} = \prob((B_j, E_j) \in \beta_i \mid \vec{\theta}, i(j) = i)$.
\end{itemize}

\begin{figure}
\includegraphics[width=\textwidth]{cis-perfect-testing/regions_diag}
\caption[Admissible, inadmissible, and undetected regions]{%
  The regions for individual $i$.
  $i$'s first test was at time 0 and negative, with subsequent negative tests at times 7, 14, and 56, and positive tests at times 21 and 28.
  Inadmissible regions are unshaded.
  The black (impossible) region covers the cases in which recovery occurs prior to infection.
  \label{perf-test:fig:partitionSpace}
}
\end{figure}

@@

Each episode occurs in one of the classes independently and with the same probability.
Therefore, conditional on $\ntot$, the number of episodes in each class is multinomial.
Specifically:
\begin{align}
  \begin{bmatrix}
    n_{1a} \\ n_{1u} \\ n_{1\bar{a}} \\ \vdots \\ n_{\Ncis,a} \\ n_{\Ncis,u} \\ n_{\Ncis,\bar{a}}
  \end{bmatrix}
  \mid \ntot
  =
  \MN\left(
    \ntot,
    \frac{1}{\Ncis}
    \begin{bmatrix}
      p_{1a} \\ p_{1u} \\ p_{1\bar{a}} \\ \vdots \\ p_{\Ncis,a} \\ p_{\Ncis,u} \\ p_{\Ncis,\bar{a}}
    \end{bmatrix}
  \right)
  \label{perf-test:eq:multinomial}
\end{align}

@@

\subsection{Posterior density} \label{perf-test:sec:posterior-density}

Let $\na = (n_{1a}, \dots, n_{\Ncis a})^T$ be the vector of counts of episodes in the admissible classes.
This represents all the observations used in \cref{perf-test:eq:multinomial}.
Therefore, the posterior density we are interested in is $p(\vec{\theta} \mid \na)$.
It can be written as follows.
\begin{align}
p(\vec{\theta} \mid \na)
&\propto p(\vec{\theta}) p(\na \mid \vec{\theta}) \\
&= p(\vec{\theta}) \sum_{\ntot=\ndet}^\infty p(\ntot \mid \vec{\theta}) p(\na \mid \ntot, \vec{\theta})
\label{perf-test:eq:posterior1}
\end{align}
where the last line augments the likelihood with the total number of episodes, $\ntot$.

First, I derive the form of $p(\na \mid \ntot, \vec{\theta})$.
From \cref{perf-test:sec:episode-types}, the counts $\na$ and $\nnodet$'s joint distribution, conditional on $\ntot$, is multinomial.
All $n_{i\bar{a}} = 0$, therefore these classes can be ignored; counts of zero do not contribute to the multinomial probability mass function (they multiply by 1).
The support of the distribution requires $\nnodet = \ntot - \sum_{i=1}^{\Ncis} n_{ia} = \ntot - \ndet$ and $\ntot \geq \ndet$.
Hence:
\begin{align}
p(\na \mid \ntot, \vec{\theta})
&= p(\nnodet  = \ntot - \na, \na\mid \ntot, \vec{\theta}) \\
&= \frac{\ntot!}{(\ntot - \ndet)! \prod_{i=1}^{\Ncis} n_{ia}!}  \pnodet^{\ntot - \ndet}\prod_{i=1}^{\Ncis} \left(\frac{1}{\Ncis}p_{ia}\right)^{n_{ia}} \\
&\propto \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \label{perf-test:eq:augmented-likelihood}
\end{align}
the final line follows because $n_{ia} = 1$ for $i \in \set{D}$ and $n_{ia} = 0$ otherwise.

Substituting \cref{perf-test:eq:augmented-likelihood} into \cref{perf-test:eq:posterior1}:
\begin{align}
&\propto p(\vec{\theta}) \sum_{\ntot=\ndet}^\infty \left( p(\ntot \mid \vec{\theta}) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \prod_{i \in \set{D}} p_{ia} \right) \\
&= p(\vec{\theta}) \left( \prod_{i \in \set{D}} p_{ia} \right) \left( \sum_{\ntot=\ndet}^\infty p(\ntot \mid \vec{\theta}) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet} \right) \label{perf-test:eq:posterior2}.
\end{align}
For convenience, define the summation term as:
\begin{align}
  \eta =  \sum_{\ntot=\ndet}^\infty p(\ntot \mid \vec{\theta}) \frac{\ntot!}{(\ntot - \ndet)!} \pnodet^{\ntot - \ndet}.
  \label{perf-test:eq:eta}
\end{align}
The rest of this section derives expressions for each of $p_{ia}$, $p_{u}$ and $\eta$.

\subsection{Expression for $\eta$}

I derive an analytical solution to $\eta$ under the prior $\ntot \dist \NBc(\mu, r)$.
I assume that $\vec{\theta}$, which are only the parameters of the survival distribution, has a prior independent of $\ntot$, the number of infections that occurred in the cohort.
Therefore, $p(\ntot \mid \vec{\theta}) = p(\ntot)$.

Putting a negative binomial prior on $\ntot$ is equivalent to the following gamma-Poisson composite; its use simplifies the derivation.
\begin{align}
\ntot \mid \lambda &\dist \Poi(\lambda) \\
\lambda &\dist \GamDist(a, b)
\end{align}
where $b = r / \mu$ and $a = r$.
Hence:
\begin{align}
\eta
&= \int \sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} p(\ntot \mid \lambda) p(\lambda) d\lambda &\text{$\lambda$ explicit}\\
&= \int \sum_{\ntot=\ndet}^\infty \frac{\ntot!}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} \frac{\lambda^{\ntot} e^{-\lambda}}{\ntot!} p(\lambda) d\lambda &\ntot \dist \Poi\\
%&= \int \sum_{\ntot=\ndet}^\infty \frac{1}{(\ntot-\ndet)!} \pnodet^{\ntot-\ndet} \lambda^{\ntot-\ndet} \lambda^{\ndet} e^{-\lambda} p(\lambda) d\lambda \\
&= \int \lambda^{\ndet} e^{-\lambda} p(\lambda) \sum_{\nnodet=0}^\infty \frac{(\pnodet \lambda)^{\nnodet}}{\nnodet!} d\lambda &\nnodet = \ntot-\ndet\\
&= \int \lambda^{\ndet} e^{-\lambda} p(\lambda) e^{\lambda \pnodet} d\lambda &\text{Maclaurin series of $e$} \\
&= \int \lambda^{\ndet} e^{-\lambda(1 - \pnodet)} \frac{b^a}{\Gamma(a)} \lambda^{a-1} e^{-b\lambda} \lambda d\lambda &\text{Gamma pdf}\\
&= \int \frac{b^a}{\Gamma(a)} \lambda^{a+\ndet-1} e^{-(b+1-\pnodet)\lambda} \lambda d\lambda \\
&= \frac{b^a}{\Gamma(a)} \frac{\Gamma(a+\ndet)}{(b+1-\pnodet)^{a+\ndet}} &\text{Gamma pdf}\\
&\propto (b+1-\pnodet)^{-(a+\ndet)} \\
&= (r/\mu + 1 - \pnodet)^{-(r+\ndet)} &\text{sub in $\mu$ and $r$}\\
&\propto(r + \mu (1- \pnodet))^{-(r+\ndet)}.
\end{align}

@@

An episode $j$ in individual $i(j)$ is detected if and only if all the following conditions are met.
\begin{enumerate}
    \item $\exists t \in \sched_{i(j)}$ such that $b_j \leq t \leq e_j$; this condition is equivalent to having at least one positive test for the episode.
    \item $\exists t \in \sched_{i(j)}$ such that $t < b_j$.
      For individuals enrolled during the period considered ($\min \sched_{i(j)} > 0$), this ensures that the beginning of the episode is lower bounded; thereby adjusting for those who enrolled during the period being less likely to have a detected infection.
      For individuals enrolled prior to the period considered ($\min \sched_{i(j)} \leq 0$), this means that the episode was not detected prior to time 1.
    \item $\exists t \in \sched_{i(j)}$ such that $t \geq b_j$.
      This means that the test is detected before $T$.
    \item $\exists t \in \sched_{i(j)}$ such that $t > e_j$, upper bounding the end of the episode.
      For episodes detected in the period I consider, this happens a negligible number of times, only in individuals lost to follow-up; therefore, I neglect this possibility.
      For a new context, including recent infections, this condition could be relaxed by considering episodes that do not meet this criteria as right censored.
\end{enumerate}

% An episode is undetected if and only if no tests are performed during the episode or if there was no negative test prior to the episode.
% Equivalently, that the first test at or after $b$ is after $e$, or that there is no negative test prior to $b$.
To formalise this, first define $\tau_{\sched_i}(t)$ as the time until the next test at or after time $t$ in the schedule $\sched_i$:
\begin{align}
\tau_{\sched_i}(t) &= \min \{ t' \in \sched_i : t' \geq t \} - t
\end{align}
defining $\min \emptyset = \infty$; \ie $\tau_{\sched_i}(t) = \infty$ if there is no $t' \in \sched_i$ such that $t' \geq t$.
The condition that $\exists t \in \sched_{i(j)}$ such that $b_j \leq t \leq e_j$ can now be expressed as $e_j \geq b_j + \tau_{\sched_{i(j)}}(b_j)$.
Equivalently, $d_j \geq \tau_{\sched_{i(j)}}(b_j) + 1$.

% Then $\Omega_i$ can be written as:
% \begin{align}
% \Omega_i = \{ (b, e) \ssep \tau_{\sched_i}(b) + b > e \vee b \leq \min(\sched_i) \}.
% \end{align}
Therefore, omitting the conditioning on $\vec{\theta}$ and $i(j) = i$:
\begin{align}
1 - p_{iu}
&= 1 - \prob((B_{j}, E_{j}) \in \Omega_i) \\
&= \prob(D_j \geq \tau_{\sched_{i}}(B_j)+ 1, B_j > \min \sched_{j}) \\
&= \sum_{b = \min \sched_{i} + 1}^T \prob(D_j \geq \tau_{\sched_{i}}(b) + 1 \mid B_j = b) \prob(B_j = b)\\
&= \frac{1}{T} \sum_{b = \min \sched_{i} + 1}^T S_{\vec{\theta}}(\tau_{\sched_{i}}(b) + 1)
\end{align}
defining $S_{\vec{\theta}}(\infty) = 0$.

Substituting into \cref{perf-test:eq:pu}:
\begin{align}
1 - p_u
& \propto \sum_{i=1}^{\Ncis} \sum_{b = \min \sched_{i} + 1}^T S_{\vec{\theta}}(\tau_{\sched_{i}}(b) + 1)
\end{align}
